---
title: "Knapsack package"
author: "Pedram , Mariano and Roshni "
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
  #code_folding: show
vignette: >
  %\VignetteIndexEntry{Knapsack package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette summarizes three functions of Knapsack package and gives examples on how to use them.
The 3 functions are,

1. Brute force algorithm
2. Dynamic algorithm
3. Greedy approximation algorithm

These functions takes a data.frame x with two variables v and w and returns the maximum knapsack value and which elements (rows in the data.frame).

### 1.Brute force algorithm

```{r}
brute_force_knapsack <- function(x,W)
{
  stopifnot(is.data.frame(x))
  stopifnot(W>0)
  value <- 0
  elem <-c()
  
  for(i in 1:(2^nrow(x)-1))
  {wsum <- 0
  vsum <- 0
  loop <-c()
  binary <- intToBits(i)
  
  for(j in 1:length(binary))
  {if( binary[j] == TRUE )
  {
    wsum <- wsum + x[j,1]
    vsum <- vsum + x[j,2]
    loop <- c(loop,j)
  }
    
  }
  
  if(vsum > value && wsum <= W)
  {value<-vsum 
  elem<-loop}}
  
  return(list(value=round(unname(value)),elements=elem))
}
```


### Paralellized version
```{r}

library(parallel)
par <- function(x,W)
{
  last_value <- 0
  last_vector <-c()
  rows<-(2^nrow(x)-1) * 32
  
  #first int function, it receives all combinations (7x32)
  step1 <- function(rows)
  {
    w_sum <- 0
    v_sum <- 0
    last_p <-0
    i <- 1:(2^nrow(x)-1)
    j <- 1:32
    #matrix repeats 32 (7 times), repeat 7 (32 times)
    mat_ij <- matrix(c(rep ( j , (2^nrow(x)-1) ),rep (i,32)),ncol=2)
    binary <- intToBits(mat_ij[rows,2])
    #for 1 (1), for 2 (2), for 3 (1,2) for 4 (3), for 5(1,3)...
    if (binary[mat_ij[rows,1]] == TRUE)
    {
      w_sum <- w_sum + unname(x[mat_ij[rows,1],1])
      v_sum <- v_sum + unname(x[mat_ij[rows,1],2])
      last_p <- mat_ij[rows,1]
    }
    list(w=w_sum,v=v_sum,p=last_p)
  }
  
  all_posibilities <-mclapply(1:rows,step1, mc.cores = 2)
  resvec <- unlist(all_posibilities) 
  
  
  
  step2 <- function(i)
  {
    start <- 96*(i-1)+1
    end <- 96*i
    temp1 <- resvec[start:end]
    weightsum <- sum(temp1[which(names(temp1) == "w")])
    weightsum <- unname(weightsum)
    valuesum <- sum(temp1[which(names(temp1) == "v")])
    valuesum <-unname(valuesum)
    packetschosen <- temp1[which(names(temp1) == "p")]
    packetschosen <- packetschosen[-which(packetschosen == 0)]
    packetschosen <-unname(packetschosen)
    list(w=weightsum,v=valuesum,p=packetschosen)
  }
  #only considering valid options.
  list_valid_weight <-mclapply(1:(rows * 3 / 96),step2, mc.cores = 2)
  
  
  lengthy <-length(list_valid_weight)
  
  step3 <- function(number){
    if(list_valid_weight[[number]]$w > W){
      return(list(v=0,p=c(0,0)))
    }else{
      return(list(v=unname(list_valid_weight[[number]]$v),p=unname(list_valid_weight[[number]]$p)))
    }
  }
  
  step3 <-mclapply(1:lengthy,step3, mc.cores = 2)
  
  values <- unlist(step3)
  values <- values[which(names(values)=="v")]
  last_value <- max(values)
  last_vector <- max.col(matrix(values,nrow=1)) 
  list(value=round(unname(last_value)),elements = step3[[last_vector]]$p)
}
```
#doesnt work on windows stackoverflow.com/questions/17196261/understanding-the-differences-between-mclapply-and-parlapply-in-r


#### Run time for 16 objects

```{r}
ptm <- proc.time()
n <- 16
knapsack_objects <- data.frame(
  w=sample(1:4000, size = n, replace = TRUE),
  v=runif(n = n, 0, 10000))
brute_force_knapsack(x = knapsack_objects[1:12,], W = 2000)
proc.time() - ptm
```

### 2.Dynamic algorithm

```{r}
knapsack_dynamic <- function(x, W){
  stopifnot(is.data.frame(x) == TRUE)
  stopifnot(is.numeric(W) == TRUE)
  n <- nrow(x)
  w <- x[[1]]  #weights
  v <- x[[2]]  #values
  capacity <- W     #capacity of knapsack
  elem <- c()
  m <- matrix(0,nrow = n+1,ncol = capacity+1)
 
  #to return the max value 
  for (i in 2:n){
    for (j in 1:capacity){
      if (j > w[i]) 
        m[i, j] <- max(m[i-1, j-w[i]] + v[i], m[i-1, j])
      else 
        m[i,j] <- m[i-1, j]
    }
  }
  
  val <- m[i, j]
  value <- round(val)
  
  #to print the elements
  while(capacity>0 && n-1>0){
    if(m[n,capacity]!=m[n-1,capacity]){
      elem<-c(elem,n)
      capacity<-capacity-w[n]
      elements <- sort(elem)
    }
    n<-n-1
  }
  return(list("value"=value,"elements"=elements))
  
}
```

#### Run time for 500 objects

```{r}
ptm <- proc.time()
n <- 500
knapsack_objects <- data.frame(
  w=sample(1:4000, size = n, replace = TRUE),
  v=runif(n = n, 0, 10000))
knapsack_dynamic(x = knapsack_objects[1:12,], W = 2000)
proc.time() - ptm
```

### 3.Greedy approximation algorithm

```{r}
greedy_knapsack <- function(x,W){
  stopifnot(is.data.frame(x))
  stopifnot(W>0)
  df <- transform(x, c= v/w )
  df<- df[order(df$c,decreasing = TRUE),]
  w8  <- vector()
  j=0
  Val <- vector()
  elements <- vector()
  while(sum(w8) <= W)
  {
    w8 <-  append(w8,df$w[j])
    Val <- append(Val,df$v[j])
    j <- j+1
    
  }
  
  if(sum(w8)> W)
  { 
    w8 <- head(w8,-1)
    Val <- head(Val,-1)
  }
  Val
  elements <- append(elements,which(x$w  %in% w8))
  results <- list("value"= sum(Val),"elements"=elements)
  return(results)
  
}
```

#### Run time for 1000000 objects

```{r}
ptm <- proc.time()
n <- 1000000
knapsack_objects <- data.frame(
  w=sample(1:4000, size = n, replace = TRUE),
  v=runif(n = n, 0, 10000))
greedy_knapsack(x = knapsack_objects[1:1200,], W = 2000)
proc.time() - ptm
```

### Examples

#### For brute force method

```{r}
#brute_force_knapsack(x = knapsack_objects[1:8,], W = 3500)
#brute_force_knapsack(x = knapsack_objects[1:12,], W = 2000)
```


#### For dynamic approach

```{r}
#knapsack_dynamic(x = knapsack_objects[1:8,], W = 3500)
#knapsack_dynamic(x = knapsack_objects[1:12,], W = 2000)
```

#### For greedy approach
```{r}
#greedy_knapsack(x = knapsack_objects[1:800,], W = 3500)
#greedy_knapsack(x = knapsack_objects[1:1200,], W = 2000)
```


### See also

1. "Description of Knapsack problem"
([via](https://en.wikipedia.org/wiki/Knapsack_problem))

2. "Pseudocode to implement dynamic algorithm"
([via](https://en.wikipedia.org/wiki/Knapsack_problem#0.2F1_knapsack_problem))

3. "To implement greedy approach"
([via](https://en.wikipedia.org/wiki/Knapsack_problem#Greedy_approximation_algorithm))

4. "To measure the run time of code chunk"
([via](https://www.r-bloggers.com/5-ways-to-measure-running-time-of-r-code/))
